Full Brain Blood-Oxygen-Level-Dependent Signal Parameter Estimation Using Particle Filters

### 2010-08-23 doc/paper/main.tex
updated paper

\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{array}

@@

%\newcommand{\gettitle}{\text{Full Brain Blood-Oxygen-Level-Dependent Signal Parameter Estimation Using Particle Filters}}

@@

\author{{Micah~Chambers}% <-this % stops a space
\thanks{M. Chambers is with the Department
of Electrical and Computer Engineering, Virginia Tech University, Blacksburg,
VA, 24060 USA e-mail: (micahc@vt.edu}% <-this % stops a space
\thanks{Manuscript received April 19, 2005; revised January 11, 2007.}}

\markboth{Journal of \LaTeX\ Class Files,~Vol.~6, No.~1, January~2007}%
{Chambers: Full Brain Blood-Oxygen-Level-Dependent Signal Parameter Estimation Using Particle Filters}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

@@

To prevent transients at the beginning of the FMRI run, the first 9
measurements (18.9 seconds) were dropped.
One knot was used for every 20 time points (at a TR of 2.1), with the first
knot and last knot taking half as many points as the rest of the knots.
The median level of each group of measurements was calculated, and the
knot placed at the average time of the group. After subtracting this spline
from the timeseries, the resulting timeseries was divided by the original
sample mean. To compensate for the BOLD signal not generally being zero-mean,
the Median-Absolute-Deviation was added to the final signal (\autoref{eq:mad}). 
\begin{equation}
y_{\text{gain}, 0:K} = 1.4826\underset{i=0:K}{\text{median}}(y_i - \text{median}(y_{0:K}))
\label{eq:mad}
\end{equation}

@@

\subsection{Data Acquisition}
\label{sec:MethodsData}

\section{Results}
\label{sec:Results}
\subsection{Simulation}
The final
histogram for a single active region, with a single set of true 
parameters is shown in \autoref{fig:slicesim_hist2}.

@@

\begin{itemize}
\item Long Single Result: Correlation Matrix, error in estimate
\item Possum Results: Histogram of parameters for a region
\item FMRI Results: Activation maps comparing M.I. with SPM

>>>>

\usepackage{colortbl}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{array}

@@

@@

%TODO change manuscript recieved/revised
\author{{Micah~Chambers}% <-this % stops a space
\thanks{M. Chambers is with the Department
of Electrical and Computer Engineering, Virginia Tech University, Blacksburg,
VA, 24060 USA e-mail: (micahc@vt.edu}% <-this % stops a space
\thanks{Manuscript received April 19, 2005; revised January 11, 2007.}}

%TODO change journal name
\markboth{Journal of ...}%
{Chambers: Full Brain Blood-Oxygen-Level-Dependent Signal Parameter Estimation Using Particle Filters}
% The only time the second header will appear is for the odd numbered pages
% after the title page when using the twoside option.
% 
% *** Note that you probably will NOT want to include the author's ***
% *** name in the headers of peer review papers.                   ***
% You can use \ifCLASSOPTIONpeerreview for conditional compilation here if
% you desire.

@@

To prevent transients at the beginning of the FMRI run, the first 9
measurements (18.9 seconds) were dropped. Each of the remaining 
images were registered to the new first image, to account for
movement. 
One knot was used for every 20 time points (at a TR of 2.1), with the first
knot and last knot taking half as many points as the rest of the knots.
The median level of each group of measurements was calculated, and the
knot placed at the average time of the group. After subtracting this spline
from the timeseries, the resulting timeseries was divided by the original
sample mean. To compensate for the BOLD signal not generally being zero-mean,
the Median-Absolute-Deviation was added to the final signal (\autoref{eq:mad}). 
\begin{equation}
y_{\text{gain}, 0:K} = 1.4826\underset{i=0:K}{\text{median}}(y_i - \text{median}(y_{0:K}))
\label{eq:mad}
\end{equation}

For SPM, the first 9 measurements were again dropped, and the images
co-registered across time. The co-registered data was spatially smoothed,
and then filtered with an adaptive cut off (which is built into the
SPM analysis). For the SPM analysis, the Canonical HRF was used
and model time derivatives were included (but not in the contrast
vector). All other settings used in SPM8 were left at the default
for analyzing FMRI data.

@@

\subsection{Data Acquisition}
\label{sec:MethodsData}
For the FMRI data discussed in \autoref{sec:RealData}, tests were
performed on a right handed volunteer using a GE SIGNA HDx 1.5 Tesla
scanner with a single echo EPI sequence. Slice spacing was
5mm, and pixel sizes were 3.75mm.  Repetition Time was 2.1s,
Echo Time was 40 ms and the imaging frequency was set to $63.854$MHz.
The image resolution was $64 \times 64 \times 28$.
The subject was presented with either a single or
double flash and was asked to respond with a right handed
or left handed finger tap, respectively. The FMRI began 18.9 seconds
before the beginning of the experiment, to allow for transients
in to image to settle out. The timing of the flashes are shown
below with time 0 corresponding to the beginning of
the 10th TR (so 9 images were dropped).

The timing of the single flashes were:
1.706, 11.944, 17.063, 18.769, 34.125, 39.244, 44.231, 47.644, 49.350,
61.294, 64.706, 66.413, 69.825, 71.531, 73.238, 76.650, 80.063,
90.169, 96.994, 110.644, 117.469, 120.881, 130.988, 132.694, 154.875,
158.288, 161.700, 165.113, 168.525, 176.925, 178.631, 183.750, 190.575,
204.225, 205.931, 211.050, 216.038, 222.863, 226.275, 236.513, 248.456,
255.281, 258.563, 263.681, 273.919, 277.331, 287.569, 292.688, 294.394,
299.381

Timing of the double flashes were:
0.131, 6.825, 20.475, 27.300, 35.831, 52.763, 54.469, 59.588, 86.756,
91.875, 107.231, 108.938, 112.350, 114.056, 115.763, 119.175, 126.000,
134.400, 136.106, 141.225, 144.638, 156.581, 159.994, 166.819, 171.806,
175.219, 185.456, 188.869, 202.519, 212.756, 217.744, 221.156, 227.981,
229.688, 233.100, 243.338, 245.044, 246.750, 250.163, 261.975, 270.506,
272.213, 280.744, 282.450, 289.275, 296.100, 301.088, 304.500

\section{Results}
\label{sec:Results}
\subsection{Simulation}
\begin{table*}[t]
\centering
\begin{tabular}{|c | c  c  c  c  c  c  |}
\hline
  & $\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$ \\
\hline
$\alpha$                      & 0.889884 & & & & & \\
\rowcolor[gray]{.8} $E_0$     & -0.7661395 & -0.5230723 & & & & \\
$V_0$                         & 0.6244049 & 0.4239271 & -0.7964774 & & & \\
\rowcolor[gray]{.8} $\tau_s$  & 0.6204843 & 0.295425 & -0.7481253 & 0.3440421 & & \\
$\tau_f$                      & -0.0004259 & -0.3966881 & -0.4314174 & 0.1962954 & 0.6990775 & \\
\rowcolor[gray]{.8} $\epsilon$& 0.6158116 & 0.6558179 & -0.641348 & 0.2846632 & 0.4458142 & -0.097079 \\
\hline
\end{tabular}
\caption{Correlation of parameter estimates for a 2400 second simulation.}
\label{tab:long_corr}
\end{table*}

For a 2400 second simulated FMRI run, a signal was generated using the 
following parameters:
$\{\tau_0 = 1.45, \alpha = 0.3, E_0 = 0.47, V_0 = 0.044, \tau_s = 1.94, \tau_f = 1.99, \epsilon = 1.8\}$,
with measurement noise ($\sigma_y$) of $0.001$ and drift standard deviation
($\sigma_x$) of $0.0005$. The minimum and maximum error particles had MSE of $ $ and $ $,
respectively, while the mean of the particles gave MSE of $ $. %TODO fill in errors
The correlation matrix for the final parameter estimates is shown in \autoref{tab:long_corr}.
The significant correlation of parameters indicate that no one estimate of 
particles dominate the solution. Additionally, given the very low error in the BOLD,
the energy graph is full of global minima. The final histogram for a single active region, 
with a single set of true parameters is shown in \autoref{fig:slicesim_hist2}.
This also indicates that ill-defined nature of the parameters.

@@

\begin{figure}[!t]
\subfloat[Region labels for simulated slice.]
{\includegraphics[width=2.5in]{images/simregions.png}\label{fig:simslice_mask}}
\hfil
\subfloat[Mutual Information Map. Scale is bits. Higher (yellow) is better.]
{\label{fig:sim_hm_mi} \includegraphics[width=2.5in]{images/sim_hm_mi}}
\caption{Comparison of activation with greymatter, parameter regions.}
\label{fig:simslice_hm}
\end{figure}

\begin{itemize}
\item Long Single Result: Correlation Matrix, error in estimate
\item Possum Results: Histogram of parameters for a region
\item FMRI Results: Activation maps comparing M.I. with SPM
\end{itemize}
